### 2024-07-28 manuscript/auxiliary/figure-styles.tex
Adding dependencies

%------------------------------------- TikZ figure styles ----------------------
\usepackage{amssymb}
\usepackage{tikz}
\usetikzlibrary{tikzmark}
\usetikzlibrary{shapes,arrows,arrows.meta,positioning}
\usetikzlibrary{shapes.multipart}

\newcommand{\MainlineShiftX}{0mm}
\newcommand{\AssignInFig}[2]{\makebox[5mm][l]{\ensuremath{#1}}\ensuremath{\gets#2}}

% Boundary / labeling nodes
\tikzstyle{boxed} = [rectangle, draw, dashed, minimum height=12.50cm, minimum width=19.25cm]

\tikzstyle{timebox} = [rectangle, draw, dashed, minimum height=5.7cm, minimum width=1.35cm]

\tikzstyle{label} = [text width=9.5cm, text centered,minimum width=9.5cm, rotate=90 ]

% Actor nodes	
\tikzstyle{ident} = [circle, draw, fill=green!40, text width=2em, text centered, minimum height=2em, node distance=3cm]

\tikzstyle{invis} = [text width=0em, text centered, minimum height=0em, node distance=3cm]

\tikzstyle{ether} = [text width=0em, text centered, minimum height=0em, node distance=0cm]

\tikzstyle{agent} = [ellipse, draw, fill=green!40, text width=6em, text centered, minimum height=2em, node distance=2cm]

\tikzstyle{flow} = [chamfered rectangle, chamfered rectangle xsep=2cm, chamfered rectangle angle=30, draw, fill=yellow!30, text width=13mm, text centered, minimum height=2em, node distance=1cm]

\tikzstyle{godo} = [regular polygon, regular polygon sides=6, shape aspect=0.5, minimum width=1cm, minimum height=0.75cm, draw, fill=yellow!30, text width=1em, text centered, minimum height=2em, node distance=0.25cm]

\tikzstyle{flip} = [chamfered rectangle, chamfered rectangle xsep=2cm, chamfered rectangle angle=30, draw, fill=yellow!30, text width=13mm, text centered, minimum height=2em, node distance=1cm]

\tikzstyle{flipBig} = [chamfered rectangle, chamfered rectangle xsep=2cm, chamfered rectangle angle=30, draw, inner sep=1mm, fill=yellow!30, text width=12mm, text centered, minimum height=2.5em, node distance=1cm]

\tikzstyle{which} = [chamfered rectangle, chamfered rectangle xsep=2cm, chamfered rectangle angle=30, draw, fill=yellow!30, text width=13mm, text centered, minimum height=2em, node distance=1cm ]

\tikzstyle{flowBig} = [chamfered rectangle, chamfered rectangle xsep=2cm, chamfered rectangle angle=30, draw, fill=yellow!30, text width=18mm, text centered, minimum height=4em, node distance=1cm]

\tikzstyle{flowMed} = [chamfered rectangle, chamfered rectangle xsep=3cm, chamfered rectangle angle=30, draw, fill=yellow!30, text width=30mm, text centered, minimum height=2em, node distance=1cm]

% Game/protocol interfaces
\tikzstyle{oracl} = [rectangle, draw, fill=blue!15, text width=40mm, text centered, minimum height=2em, node distance=3cm]

\tikzstyle{algor} = [diamond, draw, fill=blue!30, text width=3.75em, text badly centered, node distance=2.5cm, inner sep=0pt]

\tikzstyle{cntrl} = [regular polygon, regular polygon sides=6, draw, fill=red!35, text width=1.2cm, text centered, minimum height=1.4cm, node distance=1cm, inner sep=0.0625mm]

\tikzstyle{ended} = [regular polygon, regular polygon sides=8, draw, fill=red!35, text width=1.2cm, text centered, minimum height=1.4cm, node distance=1cm, inner sep=0.0625mm]

\tikzstyle{timeline} = [rectangle, draw, fill=blue!15, text width=2.5em, text centered, minimum height=2em, node distance=2cm]

% Transitions
\tikzstyle{line} = [draw, very thick, color=black!50, -latex']

\tikzstyle{arrow} = [draw, very thick, color=black!50, -{Latex[scale=1.25]}]

\tikzstyle{choice} = [draw, thick,dash pattern={on 7pt off 2pt on 1pt off 3pt}, color=black!50, -latex']

\tikzstyle{triangle} = [fill=black!50, regular polygon, regular polygon sides=3, shape border rotate=180]

% DEPRICATED!
\tikzstyle{blockOLD} = [rectangle, draw, fill=green!40, text width=10em, text centered, minimum height=4em, node distance=3cm]

\tikzstyle{cloudOLD} = [ellipse, draw, fill=yellow!60, text width=10em, text centered, node distance=3cm]

\tikzstyle{lineOLD} = [draw, thick, color=black!50, -latex']

\tikzstyle{choiceOLD} = [draw, thick,dash pattern={on 7pt off 2pt on 1pt off 3pt}, color=black!50, -latex']
%------------------------------------- TikZ figure styles ----------------------