### 2024-07-28 manuscript/figures/CGKA-Markovian.tex

\begin{tikzpicture}[ node distance = 0.5cm and 0.5cm, auto]
\node[agent, anchor=west ] (ATK) {\textbf{\((\,T, C, N\,)\) Adversary}};
\node[boxed, below left=2cm and 2cm of ATK.south west, anchor=north west] (BOXED) {};
\node[label, right=1mm of BOXED.west, anchor=north] (LABEL) {\Large \scshape Markovian, Bounded Gameplay};

\node[oracl, right=0.75cm of ATK     ] (ORC-INI) {\Oracle{init}{}};
\node[oracl, below=0.75cm of ORC-INI ] (ORC-CRT) {\Oracle{create-group}{}};
%\node[oracl, below=3.50cm of ORC-CRT  ] (ORC-DEL) {\Oracle{deliver}{}};
\node[oracl, below=4.50cm of ORC-CRT ] (ORC-COR) {\Oracle{corr}{ID_i}};
\node[oracl, below=0.50cm of ORC-COR] (ORC-HRD) {\Oracle{no-del}{ID_i}};
\node[oracl, below=0.50cm of ORC-HRD] (ORC-REV) {\Oracle{reveal}{}};
\node[oracl, below=0.50cm of ORC-REV ] (ORC-ADD) {\Oracle{add-user}{ID_i,ID_j}};
\node[oracl, below=0.50cm of ORC-ADD ] (ORC-RMV) {\Oracle{remove-user}{ID_i,ID_j}};
\node[oracl, below=0.50cm of ORC-RMV ] (ORC-UPD) {\Oracle{send-update}{ID_i}};
\node[oracl, below=3.00cm of ORC-UPD ] (ORC-CHL) {\Oracle{chall}{}};

\node[algor, right=0.50cm of ORC-INI ] (ALG-INI) {\Protocol{init}}; 
\node[algor, right=1.85cm of ORC-CRT ] (ALG-CRT) {\Protocol{create}}; 
\node[algor, right=2.75cm of ORC-ADD ] (ALG-ADD) {\Protocol{add}}; 
\node[algor, right=4.00cm of ORC-RMV ] (ALG-RMV) {\Protocol{rmv}}; 
\node[algor, right=5.25cm of ORC-UPD ] (ALG-UPD) {\Protocol{upd}};
%\node[algor, below right=0.5cm and 6.15cm of ORC-UPD ] (ALG-PRC) {\Protocol{proc}};

%\node[agent, right=4.5cm of ORC-DEL ] (DS)  {\textbf{Delivery\\Service}};
%\node[ident, right=9.25cm of ORC-HRD ] (ID1) {\textbf{$\texttt{ID}_1$}};
%\node[ident, right=9.25cm of ORC-COR ] (ID2) {\textbf{$\texttt{ID}_2$}};
%\node[invis, right=9.25cm of ORC-ADD ] (IDS) {\makebox[4mm][r]{$\vdots$}};
%\node[ident, right=9.25cm of ORC-RMV ] (IDN) {\textbf{$\texttt{ID}_N$}};
\node[invis, above left=-0.35cm and 1.90cm of ALG-CRT] (CON-CRT) {\parbox{25mm}{%
$x\in[\,2,\,N\,]$%
}};
\node[invis, above left=-0.80cm and 2.00cm of ORC-COR] (CON-COR) {\parbox{25mm}{%
If $\exists i\;\;\;$ \textsc{s.t.}\\%
$m_i\not=\bot=g_i$%
}};
\node[invis, above left=-0.80cm and 2.00cm of ORC-HRD] (CON-HRD) {\parbox{25mm}{%
If $\exists i\;\;\;$ \textsc{s.t.}\\%
$m_i\not=\bot=h_i$%
}};
\node[invis, above left=-0.25cm and 2.00cm of ORC-REV] (CON-REV) {\parbox{25mm}{%
If $c = \bot$%
}};
\node[invis, above left=-0.80cm and 2.00cm of ORC-ADD] (CON-ADD) {\parbox{25mm}{%
If $\exists i,j\,$ \textsc{s.t.}\\%
$m_i\not=\bot=m_j$%
}};
\node[invis, above left=-0.80cm and 2.00cm of ORC-RMV] (CON-RMV) {\parbox{25mm}{%
If $\exists i,j\,$ \textsc{s.t.}\\%
$m_i\not=\bot\not=m_j$%
}};

\node[flowBig, right=1.55cm of ALG-INI] (FLO-INI) {\parbox[t]{19mm}{%
\AssignInFig{g_i}{\bot},$\;\;\forall i$\\%
\AssignInFig{h_i}{\bot},$\;\;\forall i$\\%
\AssignInFig{m_i}{\bot},$\;\;\forall i$%
}};
\node[flowMed, right=0.50cm of ALG-CRT] (FLO-CRT) {\parbox[t]{32mm}{%
\AssignInFig{m_i}{\top},~$\forall i\in[\,1,\,x\,]$%
}};
\node[flow, below=1.825cm of ATK] (FLO-INC) {\parbox[t]{15mm}{%
$c\gets\bot$%
}};
\node[which, below=0.50cm of FLO-INC] (FLO-NXT) {\parbox[t]{18mm}{%
continue%
%$\!\!\!\!\triangleleft\,$If~\,$t = T$\\%
%$\phantom{\!\!\!\!\triangleleft\,}$Else$\triangledown$%
}};
\node[which, below=13.25cm of ATK] (FLO-END) {\parbox[t]{18mm}{%
$\!\!\!\!\triangleleft\,$If~\,$c \not = \bot$\\%
$\phantom{\!\!\!\!\triangleleft\,}$Else$\triangledown$%
}};

\node[invis, above left=0.75cm and 1cm of FLO-END ] (ROT-INC) {};

%\node[godo, right=0.50cm of ORC-DEL] (FLO-DEL) {\hspace*{-1mm}\sfrac{go}{do}};
%\node[godo, right=0.50cm of ORC-COR] (FLO-COR) {};
\node[flip, right=0.50cm of ORC-COR] (FLO-COR) {\AssignInFig{g_i}{\top}};
\node[flip, right=0.50cm of ORC-HRD] (FLO-HRD) {\AssignInFig{h_i}{\top}};
\node[flip, right=0.50cm of ORC-REV] (FLO-REV) {\AssignInFig{c}{\top}};
\node[flip, right=0.50cm of ORC-ADD] (FLO-ADD) {\AssignInFig{m_j}{\top}};
\node[flipBig, right=0.50cm of ORC-RMV] (FLO-RMV) {\parbox[t]{11mm}{%
\hspace*{-1.70mm}\AssignInFig{m_j}{\bot}\\%
\hspace*{-1.70mm}\AssignInFig{g_j}{\bot}%
}};
\node[flip, right=0.50cm of ORC-UPD] (FLO-UPD) {\AssignInFig{g_i}{\bot}};
\node[ended, right=1.75cm of ORC-CHL ] (RESULT) {Win\phantom{nnnnnnn}\\\phantom{mnm}Lose};

% From Adversary
\path[line  ](ATK)--(ORC-INI);
\path[line  ]([xshift=\MainlineShiftX]ATK.south)|-(ORC-CRT);
\path[line  ]([xshift=\MainlineShiftX]ATK.south)--([xshift=\MainlineShiftX]FLO-INC.north);
\path[line  ]([xshift=\MainlineShiftX]FLO-INC.south)--([xshift=\MainlineShiftX]FLO-NXT.north);

% Upper algorithms
\path[line  ](ORC-INI)--(ALG-INI);
\path[choice](ORC-CRT)--(ALG-CRT);
\draw[line  ] (ALG-INI)--(FLO-INI);
\draw[line  ] (ALG-CRT)--(FLO-CRT);

% Control-flow from CONTINUE
\path[choice]([xshift=\MainlineShiftX] FLO-NXT.south)|-(ORC-HRD);
\path[choice]([xshift=\MainlineShiftX] FLO-NXT.south)|-(ORC-COR);
\path[choice]([xshift=\MainlineShiftX] FLO-NXT.south)|-(ORC-ADD);
\path[choice]([xshift=\MainlineShiftX] FLO-NXT.south)|-(ORC-RMV);
\path[choice]([xshift=\MainlineShiftX] FLO-NXT.south)|-(ORC-UPD);
\path[choice]([xshift=\MainlineShiftX] FLO-NXT.south)|-(ORC-REV);
\path[choice]([xshift=\MainlineShiftX] FLO-NXT.south)--([xshift=\MainlineShiftX] FLO-END.north);
%\path[line  ](FLO-NXT.north west) -| ([xshift=-21mm,yshift=-7.00cm]ATK.south)|-([yshift=-2mm] ORC-CHL.west);

% Pass through to choices
\path[line  ](ORC-HRD)--(FLO-HRD.west);
\path[line  ](ORC-COR)--(FLO-COR.west);
\path[line  ](ORC-REV)--(FLO-REV);
\path[line  ](ORC-ADD)--(FLO-ADD);
\path[line  ](ORC-RMV)--(FLO-RMV);
\path[line  ](ORC-UPD)--(FLO-UPD);
\path[line  ](FLO-ADD)--(ALG-ADD);
\path[line  ](FLO-RMV)--(ALG-RMV);
\path[line  ](FLO-UPD)--(ALG-UPD);

% Branch bacward
%\path[line  ](FLO-HRD.center)--(FLO-HRD.north west);
%\path[line  ](FLO-COR.center)--(FLO-COR.north west);
%\path[line  ](FLO-ADD.center)--(FLO-ADD.south west);
%\path[line  ](FLO-RMV.center)--(FLO-RMV.south west);
%\path[line  ](FLO-UPD.center)--(FLO-UPD.south west);

%Northern Line
\path[line  ](FLO-COR.east) -| ([xshift=5mm]FLO-COR.east) |- (FLO-NXT.east);
\path[line  ](FLO-HRD.east) -| ([xshift=5mm]FLO-COR.east) |- (FLO-NXT.east);
\path[line  ](FLO-REV.east) -| ([xshift=5mm]FLO-COR.east) |- (FLO-NXT.east);
% Southern Line
\path[line  ](ALG-ADD.east) -| ([xshift=31mm] ALG-ADD.east) |- ([xshift=30mm] FLO-INC.center) |- (FLO-INC.east);
\path[line  ](ALG-RMV.east) -| ([xshift=31mm] ALG-ADD.east) |- ([xshift=30mm] FLO-INC.center) |- (FLO-INC.east);
\path[line  ](ALG-UPD.east) -| ([xshift=31mm] ALG-ADD.east) |- ([xshift=30mm] FLO-INC.center) |- (FLO-INC.east);

% From ENDGAME
\path[line](FLO-END.north west) -| ([xshift=-14mm,yshift=-5.00cm]ATK.south) |-(FLO-NXT.south west);
\path[choice]([xshift=\MainlineShiftX] FLO-END.south)|-(ORC-CHL.west);
\path[line](ORC-CHL)--(RESULT);
\draw[line width=0.375mm]([yshift=-6mm,xshift=2.5mm]RESULT.west)--([yshift=6mm,xshift=-2.5mm]RESULT.east);

\node[ether, right=1.25cm of ORC-HRD ] (PNT-HRD) {};
\node[ether, right=1.25cm of ORC-COR ] (PNT-COR) {};

%\path[line]([xshift=1.5mm,yshift=-1.5mm] ALG-ADD.north)--(DS);
%\path[line]([xshift=1.5mm,yshift=-1.5mm] ALG-RMV.north)--(DS);
%\path[line]([xshift=1.5mm,yshift=-1.5mm] ALG-UPD.north)--(DS);

%\path[choice]([xshift=6mm] DS.south)--([xshift=-0.75mm,yshift=-1.0mm] ID1.north west);
%\path[choice]([xshift=5mm] DS.south)--([xshift=-0.75mm,yshift=-1.0mm] ID2.north west);
%\path[choice]([xshift=4mm] DS.south)--([xshift= 2.00mm,yshift=-1.5mm] IDS.north west);
%\path[choice]([xshift=3mm] DS.south)--([xshift=-0.75mm,yshift=-1.0mm] IDN.north west);

%\path[choice](PNT-HRD.east)--([xshift=-1.25mm,yshift=-2.0mm] ID1.north west);
%\path[choice](PNT-HRD.east)--([xshift=-1.25mm,yshift=-2.0mm] ID2.north west);
%\path[choice](PNT-HRD.east)--([xshift=2mm    ,yshift=-2.5mm] IDS.north west);
%\path[choice](PNT-HRD.east)--([xshift=-1.25mm,yshift=-2.0mm] IDN.north west);
%
%\path[choice](PNT-COR.east)--([xshift=-1.50mm,yshift=-3.0mm] ID1.north west);
%\path[choice](PNT-COR.east)--([xshift=-1.50mm,yshift=-3.0mm] ID2.north west);
%\path[choice](PNT-COR.east)--([xshift=2mm,    yshift=-3.5mm] IDS.north west);
%\path[choice](PNT-COR.east)--([xshift=-1.50mm,yshift=-3.0mm] IDN.north west);

%\path[choice]([xshift=-1.5mm,yshift=1.5mm] ALG-ADD.north east)--([xshift=0.25mm,yshift=-1.25mm] ID1.west);
%\path[choice]([xshift=-0.5mm,yshift=0.5mm] ALG-ADD.north east)--([xshift=0.25mm,yshift=-1.25mm] ID2.west);
%\path[choice]([xshift=0.5mm,yshift=-0.5mm] ALG-ADD.north east)--([xshift=2.00mm,yshift=-1.25mm] IDS.west);
%\path[choice]([xshift=1.5mm,yshift=-1.5mm] ALG-ADD.north east)--([xshift=0.25mm,yshift=-1.25mm] IDN.west);
%
%\path[choice]([xshift=-1.5mm,yshift=1.5mm] ALG-RMV.north east)--([xshift=0.25mm,yshift=-1.25mm] ID1.west);
%\path[choice]([xshift=-0.5mm,yshift=0.5mm] ALG-RMV.north east)--([xshift=0.25mm,yshift=-1.25mm] ID2.west);
%\path[choice]([xshift=0.5mm,yshift=-0.5mm] ALG-RMV.north east)--([xshift=2.00mm,yshift=-1.25mm] IDS.west);
%\path[choice]([xshift=1.5mm,yshift=-1.5mm] ALG-RMV.north east)--([xshift=0.25mm,yshift=-1.25mm] IDN.west);
%
%\path[choice]([xshift=-1.5mm,yshift=1.5mm] ALG-UPD.north east)--([xshift=0.25mm,yshift=-1.25mm] ID1.west);
%\path[choice]([xshift=-0.5mm,yshift=0.5mm] ALG-UPD.north east)--([xshift=0.25mm,yshift=-1.25mm] ID2.west);
%\path[choice]([xshift=0.5mm,yshift=-0.5mm] ALG-UPD.north east)--([xshift=2.00mm,yshift=-1.25mm] IDS.west);
%\path[choice]([xshift=1.5mm,yshift=-1.5mm] ALG-UPD.north east)--([xshift=0.25mm,yshift=-1.25mm] IDN.west);

%\path[choice]([xshift=0.25mm,yshift=-1.25mm] ID1.west)--([xshift=-1.5mm,yshift=1.5mm] ALG-ADD.north east);
%\path[choice]([xshift=0.25mm,yshift=-1.25mm] ID2.west)--([xshift=-0.5mm,yshift=0.5mm] ALG-ADD.north east);
%\path[choice]([xshift=2.00mm,yshift=-1.25mm] IDS.west)--([xshift=0.5mm,yshift=-0.5mm] ALG-ADD.north east);
%\path[choice]([xshift=0.25mm,yshift=-1.25mm] IDN.west)--([xshift=1.5mm,yshift=-1.5mm] ALG-ADD.north east);
%
%\path[choice]([xshift=0.25mm,yshift=-1.25mm] ID1.west)--([xshift=-1.5mm,yshift=1.5mm] ALG-RMV.north east);
%\path[choice]([xshift=0.25mm,yshift=-1.25mm] ID2.west)--([xshift=-0.5mm,yshift=0.5mm] ALG-RMV.north east);
%\path[choice]([xshift=2.00mm,yshift=-1.25mm] IDS.west)--([xshift=0.5mm,yshift=-0.5mm] ALG-RMV.north east);
%\path[choice]([xshift=0.25mm,yshift=-1.25mm] IDN.west)--([xshift=1.5mm,yshift=-1.5mm] ALG-RMV.north east);
%
%\path[choice]([xshift=0.25mm,yshift=-1.25mm] ID1.west)--([xshift=-1.5mm,yshift=1.5mm] ALG-UPD.north east);
%\path[choice]([xshift=0.25mm,yshift=-1.25mm] ID2.west)--([xshift=-0.5mm,yshift=0.5mm] ALG-UPD.north east);
%\path[choice]([xshift=2.00mm,yshift=-1.25mm] IDS.west)--([xshift=0.5mm,yshift=-0.5mm] ALG-UPD.north east);
%\path[choice]([xshift=0.25mm,yshift=-1.25mm] IDN.west)--([xshift=1.5mm,yshift=-1.5mm] ALG-UPD.north east);
%
%\path[line]([xshift=-0.75mm,yshift=1mm] ID1.south west)--([xshift=-1.5mm,yshift=1.5mm] ALG-PRC.north east);
%\path[line]([xshift=-0.75mm,yshift=1mm] ID2.south west)--([xshift=-0.5mm,yshift=0.5mm] ALG-PRC.north east);
%\path[line]([xshift= 2.00mm,yshift=1mm] IDS.south west)--([xshift=0.5mm,yshift=-0.5mm] ALG-PRC.north east);
%\path[line]([xshift=-0.75mm,yshift=1mm] IDN.south west)--([xshift=1.5mm,yshift=-1.5mm] ALG-PRC.north east);

%\draw [brown] (current bounding box.south west) rectangle (current bounding box.north east);
%\draw[gray,step=0.25] (0,0) grid (20, 18);
\end{tikzpicture}