### 2024-07-28 manuscript/figures/Legend.tex

\begin{tikzpicture}[ node distance = 0.5cm and 0.5cm, auto]
\node[agent, anchor=west ] (AGT) {TreeKEM Agent};
\node[flow, right=30mm of AGT, text width=20mm] (FLO)
{\parbox{20mm}{\centering%
Control flow\\%
\&\\%
State change%
}};
\node[oracl, below=13mm of AGT, text width=28mm] (ORC) {\Oracle{\CGKAsec Oracle}{}};
\node[algor, right=30mm of ORC, text width=16mm] (ALG) {\Protocol{\parbox{12mm}{\centering\CGKAdef\\Algorithm}}};

\node[godo, below=5mm of ALG] (MUL) {};
\path[line  ](MUL.center)--(MUL.north west);
\path[line  ](ORC.south)|-(MUL.east);

\path[line  ](MUL.north west)-|(ALG.south west);
\path[line  ](MUL.east)-|(FLO.east);

\path[line](ORC)--(ALG);
\path[choice](AGT.east)--(FLO.west);
\path[choice](AGT.east)|-([xshift=-2mm, yshift=-2mm]ALG.north);

\node[invis, above left=-14mm and 27mm of FLO] (CHO) {\parbox{28mm}{%
Non-deterministic\\%
Choice%
}};

\node[invis, above left=-11mm and 32mm of ALG] (MOV) {\parbox{25mm}{%
Deterministic\\[1mm]%
Transition%
}};

\node[invis, above left=-7.5mm and 24.5mm of MUL] (FRK) {\parbox{28mm}{%
Deterministic\\%
Parallel Invocation%
}};

\end{tikzpicture}